---
title: "Setup test"
author: "Kenneth Benoit and Paul Nulty"
date: "18 October 2015"
output: html_document
---

First, you need to have **quanteda** installed.  You can do this from inside RStudio, from the Tools...Install Packages menu, or simply using
```{r, eval = FALSE}
install.packages("quanteda", dependencies = TRUE)
```

(Optional) You can install some additional corpus data from **quantedaData** using

```{r, eval=FALSE}
## the devtools package is required to install quanteda from Github
devtools::install_github("kbenoit/quantedaData")
```

Run this to test your setup.

You must have quanteda installed in order for this next step to succeed.
```{r}
require(quanteda)
```

Now summarize some texts in the Irish 2010 budget speech corpus:
```{r}
summary(ie2010Corpus)
```

Create a document-feature matrix from this corpus, removing stop words:
```{r}
ieDfm <- dfm(ie2010Corpus, ignoredFeatures = c(stopwords("english"), "will"), stem = TRUE)
```

Look at the top occuring features:
```{r}
topfeatures(ieDfm)
```

Make a word cloud:
```{r, fig.width=8, fig.height=8}
plot(ieDfm, min.freq=25, random.order=FALSE)
```

If you got this far, congratulations!







