---
title: "Text manipulation"
author: Kenneth Benoit and Paul Nulty
date: October 18th 2015
output: html_document
---

In this section we will work through some basic string manipulation functions in R.

There are several useful string manipulation functions in the R base library. In addition, we will look at the `stringr` package which provides an additional interface for simple text manipulation.

The fundamental type (or `mode`) in which R stores text is the character vector. The most simple case is a character vector of length one. The `nchar` function returns the number of characters in a character vector. 

```{r message=FALSE}
require(quanteda)
s1 <- 'my example text'
length(s1)
nchar(s1)
```

The `nchar` function is vectorized, meaning that when called on a vector it returns a value for each element of the vector.
```{r }
s2 <- c('This is', 'my example text.', 'So imaginative.')
length(s2)
nchar(s2)
sum(nchar(s2))
```

We can use this to answer some simple questions about the inaugural addresses.

Which were the longest and shortest speeches?

```{r}
which.max(nchar(inaugTexts))
which.min(nchar(inaugTexts))
```

Unlike in some other programming languages, it is not possible to index into a string in R:

```{r}
s1 <- 'This file contains many fascinating example sentences.'
s1[6:9]
```

To extract a substring, instead we use the `substr` function. 

```{r}
s1 <- 'This file contains many fascinating example sentences.'
substr(s1, 6,9)
```

Often we would like to split character vectors to extract a term of interest. This is possible using the `strsplit` function. Consider the names of the inaugural texts:

```{r}
names(inaugTexts)
# returns a list of parts
parts <- strsplit(names(inaugTexts), '-')
years <- sapply(parts, function(x) x[1])
pres <-  sapply(parts, function(x) x[2])
```

The `paste` function is used to join character vectors together. The way in which the elements are combined depends on the values of the `sep` and `collapse` arguments:

```{r}
paste('one','two','three')
paste('one','two','three', sep='_')
paste(years, pres, sep='-')
paste(years, pres, collapse='-')
```


`tolower` and `toupper` change the case of character vectors.
```{r}
tolower(s1)
toupper(s1)

```

Charcter vectors can be compared using the `==`  and `%in%` operators:
```{r}
tolower(s1) == toupper(s1)
'apples'=='oranges'
tolower(s1) == tolower(s1)
'pears' == 'pears'

c1 <- c('apples', 'oranges', 'pears')
'pears' %in% c1
c2 <- c('bananas', 'pears')
c2 %in% c1
```

The base functions for searching and replacing within text are similar to familiar commands from the other text manipulation environments, `grep` and `gsub`. The `grep` manual page provides an overview of these functions.

The `grep` command tests whether a pattern occurs within a string:

```{r}
grep('orange', 'these are oranges')
grep('pear', 'these are oranges')
grep('orange', c('apples', 'oranges', 'pears'))
grep('pears', c('apples', 'oranges', 'pears'))
```

The `gsub` command substitutes one pattern for another within a string:
```{r}
gsub('oranges', 'apples', 'these are oranges')
```

In addition to the base string operations, the `stringr` and `stringi` packages provide more extensive and more organized interfaces for string manipulation. Here we will look at some examples from the `stringr` package. You might need to install the `stringr` package using `install.packages`.

For an overview of the most frequently used functions, see the vignette: <https://cran.r-project.org/web/packages/stringr/vignettes/stringr.html>.

For an index to in-depth explanations of each of the functions, see:

```{r}
help(package='stringr')
```
